### 2016-03-17 ddays/DD16_gerasimovda.tex
Diffraction Days 2016 initial

\documentclass[12pt,twoside]{article}

\def\refname{\underline{\normalsize References}}
\def\title#1{\begin{center}{\large\bf #1}\end{center}}

\renewcommand{\author}[2]{\noindent\underbar{#1}\\#2\\}
\newcommand{\coauthor}[2]{\noindent\underbar{#1}\\#2\\}

\usepackage[paperwidth=20.5cm,paperheight=29.7cm,width=18cm,height=26.5cm,top=24mm,left=14mm,headsep=11mm]{geometry}

\usepackage{graphicx}

\usepackage{amsmath,bm}

\pagestyle{myheadings} \markboth{DAYS on DIFFRACTION 2015}{DAYS on DIFFRACTION 2015}

\begin{document}

\title{Title of paper}

\author{Author A.B., \textbf{Author B.A.}} % presenting author is marked out by \textbf 
{Institute1, address1\\ e-mail: {\tt usernameA@example.com, usernameB@example.com} }

% for name index
\index{Author, A.B.} 
\index{Author, B.A.}

\coauthor{Author C.B.} {Institute2, address2\\ e-mail: {\tt usernameC@example.com}}

\index{Author, C.B.}

%OR

%\def\supind#1{${}^\mathrm{#1}$}
%
%\author{Author A.B.\supind{1,2}, \textbf{Author B.A.}\supind{1}, Author C.B.\supind{2,3}}
%{\supind{1}Institute1, address1\\ \supind{2}Institute2, address2\\ \supind{3}Institute3,
%address3\\e-mail: {\tt usernameA@example.com, usernameB@example.com,
%usernameC@example.com} }
%
%\index{Author, A.B.} \index{Author, B.C.} \index{Author, C.B.}

Text of abstract

\medskip

\centerline{\includegraphics[width=80mm]{fig_example}}

\medskip

{\narrower\noindent \textbf{Fig.~1:} Text of long caption.\par}

\medskip

%\centerline{\textbf{Fig. 1:} Text of short caption.}

\begin{thebibliography}{9}

\bibitem{1} A. Author, B. Author, \textit{Some good book}, Some University Press, Some City
(2010).

\bibitem{2} C. Author, \textit{Journal of Something}, \textbf{90}, 34--42 (2003).

\end{thebibliography}

\end{document}