### 2016-06-05 text/parts/bundle.tex
Split into multiple files

\todo{описание}

\todo{figure}

\todo{Определитель $S$-матрицы в замкнутой форме}