### 2016-06-10 presentation/pic/resonator_bundle.tex
Some adjustments for the presentation

\newcommand{\Wglen}{6.0}; % waveguide length
\newcommand{\Warrlen}{2.0}; % wave arrow length
\newcommand{\Reslen}{4.0}; % resonator length

\coordinate (LLL) at (-\Wglen - 1, 0);
\coordinate (RRR) at ( \Wglen + 1, 0);
\coordinate (LL)  at (-\Wglen, 0);
\coordinate (RR)  at ( \Wglen, 0);
\coordinate (L)   at (-\Reslen, 0);
\coordinate (R)   at ( \Reslen, 0);
%

\coordinate (U) at (0, 3); % upper point of resonator

% waveguide
\draw[ultra thick, dotted] (LLL) -- (LL);
\draw[ultra thick] (LL)--(L);
\draw[ultra thick] (R)--(RR);
\draw[ultra thick, dotted] (RR) -- (RRR);
%

\draw[<-, thick] (-\Wglen, 1.5) -- (-\Wglen + \Warrlen, 1.5) node [midway, above] {\large $R e^{-\iu k x}$};
\draw[->, thick] (-\Wglen, 0.8) -- (-\Wglen + \Warrlen, 0.8) node [midway, above] {\large $e^{\iu k x}$};
\draw[->, thick] (\Wglen - \Warrlen, 0.5) -- (\Wglen, 0.5)   node [midway, above] {\large $T e^{\iu k x}$};

\draw (LL) node [below] {\large $\Omega_L$};
\draw (RR) node [below] {\large $\Omega_R$};

\draw[ultra thick] (0, 2) node [above] {\Large $\Omega$};

% \draw[ultra thick] node [below] {$V$} (0, 2) circle[radius=2] 
\draw[thick] (4,0)  arc[x radius = 4cm, y radius = 2cm, start angle= 0, end angle=180];
\draw[thick] (4,0)  arc[x radius = 4cm, y radius = 1cm, start angle= 0, end angle=180];
\draw[thick] (4,0)  arc[x radius = 4cm, y radius = 0.5cm, start angle= 0, end angle=180];
\draw[thick] (-4,0)  arc[x radius = 4cm, y radius = 2cm, start angle= 180, end angle=360];
\draw[thick] (-4,0)  arc[x radius = 4cm, y radius = 1cm, start angle= 180, end angle=360];
\draw[thick] (-4,0)  arc[x radius = 4cm, y radius = 0.5cm, start angle= 180, end angle=360];

\draw (-\Reslen - 0.3, 0) node [below] {\large $V_a$};
\draw ( \Reslen + 0.3, 0) node [below] {\large $V_b$};
\draw (0, 0) node {\large $\dots$};